Generating Formal Executable Sensor Models\\ from their Descriptions

### 2009-07-17 papers/BON_VHDL/paper.tex
Wrote outline for two more papers, one on VHDL<->BON refinement, and the other on SensorML<->JML refinement via BON.

%
% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass[times,10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}

@@

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}

@@

\title{The UCD CASL SenseTile System}

\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
}

@@

\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE
  Pervasive Computing}}.}

@@

This system has several novel features: (1) new deployments of \STs
requires zero configuration; (2) \STs have several unusual built-in
sensors and emitters, (3) sensor nodes can be extended via a
general-purpose USB bus, (4) \STs contain an FPGA that is dynamically
updatable on a per-experiment basis, and (5) configuring, deploying,
and managing experiments using \STs is performed by scientists via
either programmatic APIs or via a declarative graphical or textual
specification language.

@@

%======================================================================
\Section{Introduction}

Sensor networks are widely seen as an essential technology for
supporting next-generation scientific and engineering challenges,
including environmental monitoring, climate change, assisted living,
national security and intensive agriculture. Addressing these problems
requires two complementary strands of research:
\begin{enumerate}
\item to understand the \emph{specific physical and/or social
    phenomena of interest} against which to fit data collection and
  analysis; and
\item to understand the \emph{general scientific and technological
    principles governing scalable sensor networking} in order to
  maximise the leverage gained from on-going developments.
\end{enumerate}

Amongst the core general challenges are the collection of large
volumes of multimedia data, its storage, cataloging, retrieval and
processing in such a way as to minimise physical and intellectual
costs of accessing the available data. This is especially important in
pursuit of cross-disciplinary projects involving different analysis
methodologies and constraints. At the same time, it is impossible to
study sensor networking completely separate from practical
applications which provide real-world validation and verification of
the techniques being developed.

This paper describes a facility, the \ST system, to support
large-scale experiments with complex multimedia sensing and processing
at terabyte scales.  The key piece of equipment we have designed and
fabricated is a rich and reusable sensor platform, called the \ST,
that is easily deployed into the built environment for
experimentation.  This particular piece of hardware is representative
of platforms suitable for wider uses, including on-board data storage
and processing capacity and associated high-speed interconnect.

The key research challenges this system supports includes:
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
\item dealing with uncertainty in sensor systems,
\item robust management of large (multi-terabyte) sensing data on
  heterogeneous platforms,
\item interaction and activity recognition,
\item sparse signal analysis, and
\item coding and securing sensor interactions.
\end{itemize}

Underpinning each of these key challenges is a fundamental need for an
integrated, multi-modal sensor testbed that has two unique
characteristics. Firstly, it must be \emph{reconfigurable and be
  easily targeted} to collect defined data sets.  Secondly---and
perhaps most importantly---it must be \emph{of sufficient size and
  complexity} to ensure that the challenges listed above, and the
resulting theories and techniques, can be evaluated in a
scientifically rigorous manner.  It is our contention that \emph{our
  facility provides a capability---unique in the world in terms of its
  scale and flexibility---for performing in-depth investigation of
  both the specific and general issues in large-scale sensor
  networking.}  Furthermore, a synergistic benefit is the support for
the development of reference data sets and services for use by other
researchers internationally.

%=====================================================================
\Section{Conclusion}

@@

%======================================================================
%% \nocite{ex1,ex2}
\bibliographystyle{latex8}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

>>>>

%
% "Refining from the BON High-level Specification Language to the
% VHDL Hardware Description Language" for UNKNOWN09.
% Joseph R. Kiniry, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass{article} 
\usepackage{times}

@@

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}

@@

\title{Refining from the BON High-level Specification Language to the
  VHDL Hardware Description Language}

\author{Joseph R. Kiniry and Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org and dragan.stosic@gmail.com\\
}

@@

@@

The \STs have several unusual built-in sensors and emitters, sensor
nodes can be extended via a general-purpose USB bus, and \STs contain
an FPGA that is dynamically updated on a per-experiment basis.  \STs
are custom-designed in VHDL.  Like many hardware-based systems, to
write new software for or against the \ST architecture one either uses
the actual hardware or one runs a \STS.  But this simulator emulates a
piece of hardware, one would like to know if the simulator
\emph{actually} simulates the hardware.  We have developed an approach
through which we refine the VHDL specification up to the formal BON
specification language, and from this specification we generate an
executable, functional model-based specification in the JML language.
We have used this specification to implement a fully validated and
verified \ST simulator.

@@

%======================================================================
\section{Introduction}

An introduction.

%=====================================================================
\section{Conclusion}

@@

%======================================================================
%% \nocite{ex1,ex2}
\bibliographystyle{plain}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

### 2009-07-17 papers/GeneratingSensorModels/paper.tex
Wrote outline for two more papers, one on VHDL<->BON refinement, and the other on SensorML<->JML refinement via BON.

%
% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass[times,10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}

@@

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}

@@

\title{The UCD CASL SenseTile System}

@@

\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE
  Pervasive Computing}}.}

@@

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  The datastore is a
multi-terabyte scale scientific datastore into which sensor data
flows, and in which online and offline scientific computation of
sensor and other scientific data takes place.  The compute farm is a
large-scale Linux, Solaris, and OS~X-based compute farm used by
scientists for these online and offline scientific computations.

This system has several novel features: (1) new deployments of \STs
requires zero configuration; (2) \STs have several unusual built-in
sensors and emitters, (3) sensor nodes can be extended via a
general-purpose USB bus, (4) \STs contain an FPGA that is dynamically
updatable on a per-experiment basis, and (5) configuring, deploying,
and managing experiments using \STs is performed by scientists via
either programmatic APIs or via a declarative graphical or textual
specification language.

%======================================================================
\Section{Introduction}

Sensor networks are widely seen as an essential technology for
supporting next-generation scientific and engineering challenges,
including environmental monitoring, climate change, assisted living,
national security and intensive agriculture. Addressing these problems
requires two complementary strands of research:
\begin{enumerate}
\item to understand the \emph{specific physical and/or social
    phenomena of interest} against which to fit data collection and
  analysis; and
\item to understand the \emph{general scientific and technological
    principles governing scalable sensor networking} in order to
  maximise the leverage gained from on-going developments.
\end{enumerate}

Amongst the core general challenges are the collection of large
volumes of multimedia data, its storage, cataloging, retrieval and
processing in such a way as to minimise physical and intellectual
costs of accessing the available data. This is especially important in
pursuit of cross-disciplinary projects involving different analysis
methodologies and constraints. At the same time, it is impossible to
study sensor networking completely separate from practical
applications which provide real-world validation and verification of
the techniques being developed.

This paper describes a facility, the \ST system, to support
large-scale experiments with complex multimedia sensing and processing
at terabyte scales.  The key piece of equipment we have designed and
fabricated is a rich and reusable sensor platform, called the \ST,
that is easily deployed into the built environment for
experimentation.  This particular piece of hardware is representative
of platforms suitable for wider uses, including on-board data storage
and processing capacity and associated high-speed interconnect.

The key research challenges this system supports includes:
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
\item dealing with uncertainty in sensor systems,
\item robust management of large (multi-terabyte) sensing data on
  heterogeneous platforms,
\item interaction and activity recognition,
\item sparse signal analysis, and
\item coding and securing sensor interactions.
\end{itemize}

Underpinning each of these key challenges is a fundamental need for an
integrated, multi-modal sensor testbed that has two unique
characteristics. Firstly, it must be \emph{reconfigurable and be
  easily targeted} to collect defined data sets.  Secondly---and
perhaps most importantly---it must be \emph{of sufficient size and
  complexity} to ensure that the challenges listed above, and the
resulting theories and techniques, can be evaluated in a
scientifically rigorous manner.  It is our contention that \emph{our
  facility provides a capability---unique in the world in terms of its
  scale and flexibility---for performing in-depth investigation of
  both the specific and general issues in large-scale sensor
  networking.}  Furthermore, a synergistic benefit is the support for
the development of reference data sets and services for use by other
researchers internationally.

%=====================================================================
\Section{Conclusion}

>>>>

%
% "Generating Formal Executable Sensor Models from their Descriptions"
% for UNKNOWN09
% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass{article} 
\usepackage{times}

@@

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}

@@

\title{Generating Formal Executable Sensor Models\\ from their Descriptions}

@@

  In this work, rather than writing sensor simulators by hand, we
  automatically refine a SensorML specification through a series of
  stages to a formally specified, executable, functional
  model-based specification written in the Java Modeling Language.
  This resulting simulator implementation is completely validated with
  (for the most part, automatically generated) unit tests and is
  verified with the Extended Static Checker for Java, ESC/Java2.

%======================================================================
\section{Introduction}

@@

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  There are over a dozen
sensors on the \ST itself, and new sensors can be added by adding them
to the USB bus.

A software-based \STS has been designed and implemented that is meant
to mimic the \ST hardware.  A part of this simulation is, of course,
simulating the sensors themselves.  The challenge in such simulation
is to ensure that a sensor simulator properly mimics the behavior of
the actual hardware sensor.  While one can specify the behavior of
such a sensor from a hardware data sheet and then hand-implement the
sensor simulator, this process is problematic on several fronts.  Not
only is one concerned about the accuracy of the simulation, but
writing and maintaining these simulations is labor-intensive.

A popular means by which to describe sensor data is to the use the
SensorML specification language.  Using SensorML one can describe
sensors and measurement processes, including an electronic
specification sheet for a sensor, the lineage of an observation, etc.
We would hope that this declarative specification is directly related
to its simulation---in other words, the simulation must be a
refinement of the SensorML specification.  But at the moment the is no
way one can define or perform this refinement other than programming
against a SensorML API by hand.

To overcome this challenge, we: (1) formally define a sensor
meta-model in the BON specification language; (2) define a formal
refinement relation between the (XML) meta-model of SensorML and this
BON meta-model; (3) using this refinement, we automatically generate
an informal and formal BON model from any SensorML model; (4) using
the Beetlz tool, from the formal BON model, we generate a formally
specified, executable, functional model-based specification written in
the Java Modeling Language, and; (5) we validate and verify, using
unit testing and extended static checking respectively, this JML
sensor model.

Consequently, using formal BON, we can define a \emph{sensor system}
as a composition of sensor models, and thus derive a complete formally
specified and verified, executable sensor system model.  To
demonstrate this process we describe an interesting slice of the \ST
specification and its simulator.

%=====================================================================
\section{Conclusion}
